---
title: "Are ILR subcompositionally coherent?"
format: html
editor: visual
---

```{r setup, include=F}
library(compositions)
```

# Simulate

```{r}

simulate <- function(N, n_feat){
     matrix(sapply(X=1:n_feat,
                   FUN=function(x) MASS::rnegbin(n=N, mu=sample(x=5:10,size=1), theta=sample(x=1:5, size=1))),
               nrow=N, ncol=n_feat)
}

# params
N <- 100
n.features <- 100
```

# CLR

```{r}
# simulate data
data <- simulate(N, n.features)
hist(data[1,])

# clr transform
clr.data <- as.matrix(clr(data))
hist(clr.data[1,])

# remove a feature data
data.filtered <- data[,1:(n.features-1)]

# compare clr on filtered data to clr on original data
clr.data.2 <- clr(data.filtered)
portion.same <- sum(clr.data.2 == clr.data[,1:(n.features-1)])/length(clr.data.2)


print(paste0(round(portion.same*100, digits=0),
      "% of entries in the clr-transformed data are the same after removing a feature"))
```

# ILR

```{r}
# ilr transform
ilr.data <- as.matrix(ilr(data))
hist(ilr.data[1,])


# compare ilt on filtered data to ilr on original data
ilr.data.2 <- ilr(data.filtered)
n.features.ilr <- n.features - 1
portion.same <- sum(ilr.data.2 == ilr.data[,1:(n.features.ilr-1)])/length(ilr.data.2)


print(paste0(round(portion.same*100, digits=0),
      "% of entries in the ILR-transformed data are the same after removing a feature"))

```

<<<<<<< HEAD
The level of incoherence can be investigated by generating random subcompositions of the Tellus data, in this
case with zeros not replaced. For each subcomposition,
the parts were renormalized and the chi-square distances
between the parts were computed. These are compared
with the chi-square distances between the parts using the
original compositional values in the full composition. In
each case the full set of principal coordinates is computed,
thus defining the geometry of the compositional parts and
of the subcompositional parts, and then the Procrustes
correlation is used to measure the degree of concordance.

This is from the Aitchison's reappraisal paper with thom quinn.
=======
The level of incoherence can be investigated by generating random subcompositions of the Tellus data, in this case with zeros not replaced. For each subcomposition, the parts were renormalized and the chi-square distances between the parts were computed. These are compared with the chi-square distances between the parts using the original compositional values in the full composition. In each case the full set of principal coordinates is computed, thus defining the geometry of the compositional parts and of the subcompositional parts, and then the Procrustes correlation is used to measure the degree of concordance.
>>>>>>> a448c6a24002ffc34778d5d9c9f8db221e8905dd
